Ensembler: Enabling high-throughput molecular simulations at the superfamily scale

### 2015-02-11 manuscript/ms.tex
Minor edit to phrasing in abstract

\usepackage{palatino}
\usepackage{amsmath}
\usepackage{amssymb}
\pdfoutput=1
\usepackage[pdftex]{graphicx}
%\usepackage[mathbf,mathcal]{euler}
%\usepackage{citesort}
\usepackage{dcolumn}
\usepackage{boxedminipage}
\usepackage{verbatim}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}

@@

The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
However, simulations on an \emph{omics} scale are not yet widely performed, partly because software infrastructure to enable this has not kept pace. 
For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and similarities in conformations that homologous proteins can adopt.
Here, we present a new tool for enabling high-throughput simulation in the genomics era.
Ensembler takes any set of sequences---from a single sequence to an entire superfamily of interest---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement with molecular simulation to ensure stable simulation. 
The output of this pipeline is an ensemble of structures ready for subsequent parallel or distributed molecular simulations using clusters, supercomputers, or distributed computing projects like Folding@home.
Ensembler automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.  
A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
We demonstrate the power of this approach by constructing models for all catalytic domains in the human tyrosine kinase family, using all available kinase catalytic domain structures from any organism as structural templates.

>>>>

%\usepackage{palatino}

% Change to a sans serif font.
\usepackage{sourcesanspro}
\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
\usepackage[T1]{fontenc}
%\usepackage[font=sf,justification=justified]{caption}
\usepackage[font=sf]{floatrow}

% Rework captions to use sans serif font.
\makeatletter
\renewcommand\@make@capt@title[2]{%
 \@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
  {\textbf{#1}}\sf\@caption@fignum@sep#2\quad
}%
\makeatother

\usepackage{listings} % For code examples
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
\pdfoutput=1
\usepackage[pdftex]{graphicx}
%\usepackage[mathbf,mathcal]{euler}
%\usepackage{citesort}
\usepackage{dcolumn}
\usepackage{boxedminipage}
\usepackage{verbatim}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}

@@

The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
However, simulations on an \emph{omics} scale are not yet widely performed, partly because software infrastructure to enable this has not kept pace. 
For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and conformational similarities across homologous proteins.
Here, we present a new tool for enabling high-throughput simulation in the genomics era.
Ensembler takes any set of sequences---from a single sequence to an entire superfamily of interest---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement with molecular simulation to ensure stable simulation. 
The output of this pipeline is an ensemble of structures ready for subsequent parallel or distributed molecular simulations using clusters, supercomputers, or distributed computing projects like Folding@home.
Ensembler automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.  
A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
We demonstrate the power of this approach by constructing models for all catalytic domains in the human tyrosine kinase family, using all available kinase catalytic domain structures from any organism as structural templates.
